---
title: "Data Import"
output: github_document
---

## Data Import:|CVS

Let's import data using 'readr' package.

```{r}
read_csv("data_import_examples/FAS_litters.csv")
library(tidyverse)
library(readxl)
library(haven)
```

read_csv() can be used to find and read a csv document. Write relative paths. If in the same R project, you don't need to write the path in computer, just write the one in R project.

```{r}
litters_df = read_csv("data_import_examples/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
use View(litters_df) in terminal, not in rmd documents. and it will help to open a table.
Look at the data

```{r}
litters_df

head(litters_df)
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```

skim function can help us to get a quick review of the dataset.

```{r}
read_csv("data_import_examples/FAS_litters.csv", na = c("", "NA", 999, 88), skip = 1)
```
use skip

```{r}
mlb_df = read_excel("data_import_examples/mlb11.xlsx")
```

```{r}
lotr_words = 
  read_excel(
    "data_import_examples/lotR_Words.xlsx",
    range = "B3:D6"
  )
```

## Still more formats...

Read in a SAS dataset.

```{r}
pulse_df = read_sas("data_import_examples/public_pulse_data.sas7bdat")
```

## Data export

```{r}
write_csv(lotr_words, file = "results/lotr_words_df.csv")
```
write is an export function.

## Why not base r?

```{r}
dont_do_this_df = read.csv("data_import_examples/FAS_litters.csv")
```
use gitignore to ignore the data you don't want to put on GH.
