---
title: "data_manipulation with 'dplyr'"
output: html_document
date: "2022-09-22"
---
## SELECT FUNCTION
```{r}
library(tidyverse)

options(tibble.print_min = 3)

litters_data = read_csv("./data_import_examples/FAS_litters.csv",
                        col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

```
janitor::clean_names() to clean some data, for example, we clean up the column names using codes above.
col_types: c-character, d-decimal, i-integer. Insure the data is imported in the right form.

```{r}
pups_data = read_csv("./data_import_examples/FAS_pups.csv",
                     col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```
select the columns

```{r}
select(litters_data, group:gd_of_birth)
```
using":"to select a range of columns

```{r}
select(litters_data, -pups_survive)
```
using "-" to remove columns
```{r}
select(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
using "=" to rename the columns (and only select the columns you rename)
```{r}
rename(litters_data, GROUP = group, LiTtEr_NuMbEr = litter_number)
```
using rename to rename the columns (but it will show all the columns including the rename one.)

```{r}
select(litters_data, starts_with("gd"))
```

```{r}
select(litters_data, litter_number, pups_survive, everything())
```
everything will be shown, but the ones you write in front of everything() will be shown first.

```{r}
relocate(litters_data, litter_number, pups_survive)
```

## FILTER FUNCTION
Filter using comparision operators. 
*Attention: using"=="! (to distinguish from assign "=")
use %n% to detect if values appear in a set.
is.na() to find missing values
results of comparisons are logical :"TRUE" or "FALSE"
```{r}
filter(pups_data, sex == 1)
```

```{r}
filter(pups_data, sex == 2, pd_walk < 11)
```

```{r}
filter(litters_data, group %in% c("Con7", "Con8"))
```

```{r}
drop_na(litters_data)
```
the code above will omit all rows with any missing data.

```{r}
drop_na(litters_data, gd0_weight)
```
the code above will only omit the rows with missing data in gd0_weight.

## MUTATE FUNCTION
use mutate() to change columns or create new ones.
```{r}
mutate(litters_data,
       wt_gain = gd18_weight - gd0_weight,
       group = str_to_lower(group)
       )
```
create a new column called wt_gain.

## ARRANGE FUNCTION
```{r}
head(arrange(litters_data, group, pups_born_alive),10)
```
10 means only print 10 rows which having pups_born_alive data.

## %>%
piping operator "%>%" allow us to turn the nested approach into a sequential chain by passing the results of one function call as an argument to the next function call. Make it more readable. 上一行形成的数据表格下一行直接引用。
```{r}
litters_data = 
  read_csv("./data_import_examples.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  
```


