---
title: "tidy data"
output: html_document
date: "2022-09-27"
---
```{r}
library(tidyverse)
options(tibble.print_min = 5)
```

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_df
```
columns in pulse_df can be same kind of variables, such as 01_m, 06_m, ....
not tidy

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df,
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to ="bdi",
    names_prefix = "bdi_score_ "
  )

pulse_tidy_data
```
now different bdi variables are combined into one column.
```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to ="bdi",
    names_prefix = "bdi_score_ "  ) %>%
relocate(visit) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)  ) %>%
  arrange(id, visit)

print(pulse_df, n = 12)
```
arrange() let us arrange the data by id, and then by visit

```{r}
litters_wide =
  read.csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(litter_number, ends_with("weight")) %>%
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight") %>%
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))
```
recode() switch values by their name or position.

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)
```

```{r}
pivot_wider(
  analysis_result,
  names_from = "time",
  values_from = "mean")
```
groups in rows, times in columns

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")
```

```{r}
two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
  
```

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "words" ) %>%
  mutate(race = str_to_lower(race)) %>%
  select(movie, everything())
```

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(
    sex = recode(sex, '1' = "male", '2' = "female"),
    sex = factor(sex)
  )
```
should use side codes '' when '1' because it's not numeric.

```{r}
litter_data = 
  read.csv("./data/FAS_litters.csv") %>%
  janitor::clean_names()%>%
  separate(group, into = c("dose","day_of_tx"), sep = 3) %>%
  relocate(litter_number) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose)
  )
```

```{r}
fas_data = 
  left_join(pup_data, litter_data, by = "litter_number")
```
one people have several babies will be added to the dataset more than once.



